---
title: "Assignment_03"
author: "Jonathan Schuster"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  pdf_document: default
---
 <!-- This chunk is storing tidyverse, lattice, and kableExtra into my library --> 
```{r libraries, include=FALSE}
library(tidyverse)
library(lattice)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE)
```
 <!-- This chunk is styling captions on the markdown document, specifically my table captions. --> 
```{css, echo = FALSE}
caption {
      color: black;
    }
```
<!-- This chunk is loading the excel data from the internet for me. -->
```{r data, include=TRUE}
data <- read_csv(file = "https://www.zoology.ubc.ca/~whitlock/ABD/teaching/datasets/03/03e3SticklebackPlates.csv", col_types = cols())
names(data) <- c("Genotype", "Plates")


```

## Three-spined Sticklebacks
The three-spined stickleback (*Gasterosteus aculeatus*) is a fish that is native to inland coastal waters. The genus for this specific species of stickleback (*G. aculeatus*) belongs to the genus *Gasterosteus*.

<!-- This chunk is providing my markdown file with an image of the three-spined stickleback -->
```{r remedy002, fig.align='center'}
knitr::include_graphics("https://www.researchgate.net/profile/Windsor_Aguirre/publication/285915049/figure/fig2/AS:318190618988545@1452873963743/Lateral-plate-morphs-of-Gasterosteus-aculeatus-A-low-B-partial-C-complete-From_W640.jpg")

```

## Hardy Weinberg conditions
The Hardy Weinberg equation applies to this data set by showing us the adaptive divergence between the natural populations of sticklebacks. We can better understand the evolution of this specific species of sticklebacks. 

The Hardy Weinberg equation is $p^2 + 2pq + q^2 = 1$.

## Data analyses

``` {r data analyses, include = TRUE, fig.align='center', out.width = "400px"}
histogram(~ Plates | Genotype, 
		  data = data, 
		  breaks = seq(0,70,by=2),
		  layout = c(1, 3), 
		  col = "firebrick")
caption = "**Figure 2: Frequency distribution of lateral plate counts by genotype."
```

```{r data specs}
knitr::kable(data %>% group_by(Genotype) %>% summarize(N = n(), 
			 			  Mean = mean(Plates), 
			 			  Median = median(Plates), 
			 			  StdDev = sd(Plates)), 
			 align = c('l','c','c','c','c'), 
			 digits = 2,
			 format = "html", 
			 caption = "**Table 1:** Summary of lateral plate numbers by genotype.",
			 table.attr = "style='width:70%;'") %>% kable_styling(position = "center")

```
