---
title: "Assignment_03"
author: "Jonathan Schuster"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  pdf_document: default
---
 <!-- This chunk is storing tidyverse, lattice, and kableExtra into my library. Prepares my Rmd file with the necessary packages to run lines of code. --> 
```{r libraries, include=FALSE}
library(tidyverse, warn.conflicts = FALSE)
library(lattice, warn.conflicts = FALSE)
library(kableExtra, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(captioner, warn.conflicts = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```
 <!-- This chunk is styling captions on the markdown document, specifically my table captions. --> 
```{css, echo = FALSE}
caption {
      color: black;
    }
```
<!-- This chunk is loading the excel data from the internet for me. -->
```{r data, include=TRUE, message=FALSE}
data <- read_csv(file = "https://www.zoology.ubc.ca/~whitlock/ABD/teaching/datasets/03/03e3SticklebackPlates.csv")
names(data) <- c("Genotype", "Plates")


```

## Three-spined Sticklebacks
The three-spined stickleback (*Gasterosteus aculeatus*) is a fish that is native to inland coastal waters. The genus for this specific species of stickleback (*G. aculeatus*) belongs to the genus *Gasterosteus*.

<!-- This chunk is providing my markdown file with an image of the three-spined stickleback -->
```{r picture, fig.align='center', out.width = "400px", fig.cap="**Figure 1:**Lateral plate morphology of the three-spine stickleback. © Windsor E. Aguirre"}
knitr::include_graphics("https://www.researchgate.net/profile/Windsor_Aguirre/publication/285915049/figure/fig2/AS:318190618988545@1452873963743/Lateral-plate-morphs-of-Gasterosteus-aculeatus-A-low-B-partial-C-complete-From_W640.jpg")

```

## Hardy Weinberg conditions
The Hardy Weinberg equation applies to this data set by showing us the adaptive divergence between the natural populations of sticklebacks. We can better understand the evolution of this specific species of sticklebacks. 

The Hardy Weinberg equation is $p^2 + 2pq + q^2 = 1$.

## Data analyses
<!-- This chunk is provding the document with histograms presenting the stickleback research visually -->
``` {r histogram, message=FALSE, fig.align='center', fig.cap="**Figure 2:** Frequency distribution of lateral plate counts by genotype."}
histogram(~ Plates | Genotype, 
		  data = data, 
		  breaks = seq(0,70,by=2),
		  layout = c(1, 3), 
		  col = "firebrick")
```
Figure 2 is showing a table of how much percent total of sticklebacks have plates and the sizes of these plates. Table 1 summarizes the data in mm with the total numbers, mean, median, and standard deviations.
<!-- This chunk is providing the document with a table summarizing the data provided in figure 2. -->
```{r data_specs, results=FALSE}
knitr::kable(data %>% group_by(Genotype) %>% summarize(N = n(), 
			 			  Mean = mean(Plates), 
			 			  Median = median(Plates), 
			 			  StdDev = sd(Plates)), 
			 align = c('l','c','c','c','c'), 
			 digits = 2,
			 format = "html", 
			 caption = "**Table 1:** Summary of lateral plate numbers by genotype.",
			 table.attr = "style='width:70%;'") %>% kable_styling(position = "center")

```
<!-- This chunk is inserting another visual representation of the data in the form of boxplots. Looking at genotypes and plate.-->

```{r boxplots, fig.align='center', fig.cap="**Figure 3:**  Number of lateral plates by genotype."}
ggplot(data, aes(x=Genotype, y=Plates, fill="firebrick")) + 
	geom_boxplot(outlier.colour="black", outlier.size=3) + 
	ylim(0,80) + 
	guides(fill=FALSE, color=FALSE) +
	theme_bw()


```
Here we are looking at boxplots of the stickleback's and number of plates each genotype possesses. This gives us a better look at the data and shows us the outliers a bit more clearly.Moving forward, we can look in the table to get a better understanding of the differences between these genotypes through the mean, median, and standard deviations. 
<!-- In this chunk I am inserting the ANOVA table which basically gives the scientific relevance of the data for this research. -->
```{r ANOVA}
knitr::kable(anova(lm(Plates ~ Genotype, data = data)),
			 digits = 3, 
			 format = "html", 
			 caption = "**Table 2:** One-way ANOVA analysis of lateral plate numbers by genotype.",
			 table.attr = "style='width:70%;'") %>% kable_styling(position = "center")

```

## Conclusions
Through working on this project I was able to make a few conclusions. Firstly, I'll tAlk about the research provided. It looks like the sticklebacks in the middle of the sizes (mm) contained on average the most amount of lateral plates. This is due to evolutionary changes over time and selecting of genes over time.

## Reference 
Colosimo, P. F., Peichel, C. L., Nereng, K., Blackman, B. K., Shapiro, M. D., Schluter, D., & Kingsley, D. M. (2004). _The Genetic Architecture of Parallel Armor Plate Reduction in Threespine Sticklebacks._ PLoS Biology, _2(5)_, e109. https://doi.org/10.1371/journal.pbio.0020109
